ActiveCA: Time Use Data from the General Social Survey of Canada to Study Active Travel

### 2025-01-17 data-raw/01-preprocessing-episode-files.Rmd
Including labelled main files in the data paper
inclusion of section for python integration and how to use this package

@@

>>>>

Cases of episodes with `NA` weight values will be updated to value 1, meaning that since there's no weighting value for this observation, this episode only represent itself:

```{r fixing-NA-weights-walking_1992}
walking_1992$WGHT_EPI[is.na(walking_1992$WGHT_EPI)] <- 1
```

@@

Cases of episodes with `NA` weight values will be updated to value 1, meaning that since there's no weighting value for this observation, this episode only represent itself:

```{r fixing-NA-weights}
gss_episodes$WGHT_EPI[is.na(gss_episodes$WGHT_EPI)] <- 1
```

### 2025-01-17 paper/Manuscript-Data-Package.Rmd
inclusion of section for python integration and how to use this package

---
title: "ActiveCA: Time Use Data from the General Social Survey of Canada to Study Active Travel"
author:
  - name: "Anon1"
    email: "Anon1@anon.org"
    affiliation: CODE
    footnote: 1
  - name: "Anon2"
    email: "Anon2@anon.org"
    affiliation: CODE
  - name: "Anon3"
    email: "Anon3@anon.org"
    affiliation: CODE
address:
  - code: CODE
    address: "Some school"
footnote:
  - code: 1
    text: "Corresponding Author"
abstract: 
  "This paper describes {ActiveCA}, and open data product with Canadian time use data. {ActiveCA} is an `R` data package that contains analysis-ready data related to active travel spanning almost 40 years, extracted from cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), and 29 (2015) of Canada's General Social Survey. Active travel is characterized by mode, with walking being part of every cycle and bicycling starting in 1992. The attributes of active trips are the types of locations of origins and destinations, the duration of trips, and episode weights for expanding the trips to population-wide estimates. Based on the year of the survey, a variety of locations are coded. In earlier cycles, these include home, work or school, and other's home, whereas in later cycles these are augmented with locations such as grocery stores, restaurants, outdoor destinations, and others. The geographical resolution includes the province and whether the episode was in an urban or rural setting."

@@

# Original main and episode files from 2015 to display table examples
gss_m_2015 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2015/gss-m.csv"))
gss_e_2015 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2015/gss-e.csv"))

# Processed main files
data("gss_main_2015")
data("gss_main_2010")
data("gss_main_2005")
data("gss_main_1998")
data("gss_main_1992")
data("gss_main_1986")
```

@@

The main file of the time use survey compiles a large array of aggregated data, 
summarizing the answers to the questionnaire that describe households and 
individuals, as well as derived variables that summarize the respondents' use of 
time use across different activities, locations, and social interactions. This 
file documents the time and duration that respondents allocate to each activity 
and location. The Main File provides a overview of daily routines and social 
dynamics, not focusing on individual activity episodes. Additionally, this file 
categorizes activities into bigger groups and subcategories, facilitating the 
data's analytical utility with additional metrics such as total transit time, time 
spent with household members, and counts of activities and episodes.

Table \ref{tab:main-2015-unprocessed} shows the first ten rows and first six variables of the GSS PUMF 2015 main file (Cycle 29). Each row in the table correspond to a survey respondent, while the columns refer the following information: record identification (`PUMFID`), the person's weight (`WGHT_PER`), the month the survey data was collected (`SURVMNTH`), the respondent's age group (`AGEGR10`), the respondent's sex (`SEX`), and the respondent's marital status (`MARSTAT`).

The main file of the 2015 GSS surveys includes a total of 17,390 respondents, representing `r sum(gss_m_2015$WGHT_PER)` individuals and 848 variables.  For discrete variables, Statistics Canada has assigned specific codes to the possible values, with each code accompanied by a label. For instance, in the case of the variable `SURVMNTH`, a value of `1` means `January 2016`, 2 means `February 2016`, 3 corresponds to `March 2016`, and so on.

@@

```{r gss-main-file-2015}
main_2015_unprocessed <- gss_m_2015[1:10, 1:6]  %>%
  kable(
    format = "latex", 
    booktabs = TRUE, 
    longtable = TRUE, 
    align = c(rep("c", ncol(gss_m_2015[1:10, 1:6]))),
    caption = "\\label{tab:main-2015-unprocessed}Visualization of the first ten lines and first six columns of the main file of the 2015 GSS.") %>%
  kable_styling(full_width = FALSE, 
                #latex_options=c("striped", "scale_down"), 
                latex_options=c("scale_down"),
                 font_size = 8) %>%
  footnote(general = c("Legend: `PUMFID`: record identification. `WGHT_PER`:  person weight. `SURVMNTH`: survey month of data collection. `AGEGR10`: age group of the respondent. `SEX`: sex of the respondent. `MARSTAT`: marital status of the respondent."), threeparttable = T)

@@

The episode is a much bigger file that records detailed data for each activity 
episode reported by respondents. Each episode represents a single activity and its
duration, and the sum of all episodes throughout the day adds up to 24 hours. Each
entry in this file includes the start and end times of the activity, the duration,
location, and accompanying social context, informing when and where activities 
occurred and with whom. The focus of the Episode File is not on the characteristics
of the respondents, but on the characteristics of the activities, and the data 
are structured around the numerous activity instances that compose a day of the 
respondent. Although respondent-specific characteristics are not included within
the episode file, it is possible to link the main file and the episode file by 
using a key present in both the Main and Episode Files.

Similar to Table \ref{tab:main-2015-unprocessed}, which displayed an example of the main file structure, Table \ref{tab:ep-2015-unprocessed} presents the first seven episodes for the record identification number `10041` and some variables from the GSS PUMF 2015 episode file (Cycle 29). Each row in the table corresponds to an episode associated with the specified record identification (`PUMFID = 10041`), including the episode's weight (`WGHT_EPI`), episode number (`EPINO`), activity code (`TUI_01`), episode duration (`DURATION`), and episode location (`LOCATION`).

In total, the episode file of the 2015 GSS surveys contains 274,108 records, representing `r sum(gss_e_2015$WGHT_EPI)` episodes and 527 variables detailing the episodes. Similar to the main file, Statistics Canada has created codes for the discrete variables, with each value corresponding to a label.

@@

For each selected cycle of the GSS surveys, we reviewed the episode files to identify episodes of movement that involved walking or cycling. This allowed us to select the activities immediately before and after the movement episode. After that, we labeled the code variables with their appropriate descriptions, identifying each origin and destination, mode of travel, time spent in the active trip, as well the province and urban classification of the episode.

@@

For each GSS main file, we selected socioeconomic variables to help profile the individuals engaged in active trip episodes. We included key socioeconomic variables such as the respondent’s age group, sex, marital status, and number of children, among others. After this, as we did with the episode files, we labeled the code variables with their appropriate descriptions.

# How to use `ActiveCA`

This section presents some potential applications of the `ActiveCA` R package. In fact, we expect that the application of this package to extend beyond our pre-imagined range of uses.

To start the demonstration, it is first necessary to install the package. The code chunk below demonstrates how to install the `ActiveCA` package in R:

@@

After installing the `ActiveCA` R package, you can load it using the following code:

@@

For instance, if someone wants to calculate the percentage of respondents from the 2015 GSS survey with active travel episodes, it is necessary to account for the person's weights. In 2015, the weight variable is represented by `WGHT_PER.` The code chunk below demonstrates how to obtain the percentage of people with active travel episodes by age group. It uses the dplyr package to manipulate the data.

The process begins by creating a dataset that sums the population by age group. Then, it joins the 2015 episodes with the 2015 main file. Note that in both operations, the code sums the person weight variable (`WGHT_PER`) to obtain the correct population values. After this, a new dataset, `Active_percentage`, is created by merging both previous datasets. The percentage of active travel episodes by age group is calculated by dividing the total population by the population with active trip episodes, then multiplying by 100.

@@

```{r percentages-act, echo=TRUE}
# Load the dplyr library for data manipulation
library('dplyr')

@@

# Combine total and active population and calculate the percentage
Active_percentage <- Total_population %>% 
  left_join(Active_population, by = c("AGEGR10" = "AGEGR10")) %>%
  mutate(Percentage = round(Active_population/Total_population  * 100, 2))

@@

Table \ref{tab:processed-obs} displays the total number of observations processed for main and episode files. For the main files, a total of `r sum(df_main_processed$Count)` observations were processed, referring to all observations from the Time Use Surveys from 1986 to 2015, that together represents more of `r sum(df_main_processed$Weighted)` respondents (`r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '2015',]$Weighted)` for 2015, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '2010',]$Weighted)` for 2010, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '2005',]$Weighted)` for 2005, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '1998',]$Weighted)` for 1998, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '1992',]$Weighted)` for 1992, and  for 2005, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '1998',]$Weighted)` for 1998, and `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '1992',]$Weighted)` for 1986). Table X also presents the total cases of active trips episodes identified. In total `r sum(df_ep_processed$Count)` observations with register of active travel activity. Together, these observations account for `r sum(df_ep_processed$Weighted)` episodes (`r sum(df_ep_processed[df_ep_processed$Survey == 'Episodes' & df_ep_processed$Year == '2015',]$Weighted)` for 2015, `r sum(df_ep_processed[df_ep_processed$Survey == 'Episodes' & df_ep_processed$Year == '2010',]$Weighted)` for 2010, `r sum(df_ep_processed[df_ep_processed$Survey == 'Episodes' & df_ep_processed$Year == '2005',]$Weighted)` for 2005, `r sum(df_ep_processed[df_ep_processed$Survey == 'Episodes' & df_ep_processed$Year == '1998',]$Weighted)` for 1998, `r sum(df_ep_processed[df_ep_processed$Survey == 'Episodes' & df_ep_processed$Year == '1992',]$Weighted)` for 1992, and  for 2005, `r sum(df_ep_processed[df_ep_processed$Survey == 'Episodes' & df_ep_processed$Year == '1998',]$Weighted)` for 1998, and `r sum(df_ep_processed[df_ep_processed$Survey == 'Episodes' & df_ep_processed$Year == '1992',]$Weighted)` for 1986).

@@

table_df_processed <- df_processed %>%
  kable(format = "latex", booktabs = TRUE, longtable = TRUE, 
        caption = "\\label{tab:processed-obs}Total number and weighted sum of observations processed.") %>%
  kable_styling(full_width = FALSE, font_size = 10) %>%
  add_header_above(c(" " = 2, "Observations" = ncol(df_processed) - 2)) %>%
  collapse_rows(columns = 1, valign = "top", latex_hline = "major")

@@

Table \ref{tab:main-2015-processed} shows the first ten rows and first six variables of the GSS PUMF 2015 main file (Cycle 29), displayed in \ref{tab:main-2015-unprocessed} before our processing. Table \ref{tab:ep-2015-processed} presents the walking episodes for the record identification number `10041` from the GSS PUMF 2015 episode file (Cycle 29), previously displayed in Table \ref{tab:ep-2015-unprocessed}. Only active travel episodes appear in Table \ref{tab:ep-2015-processed} since the observations were filtered to select cases with walking or cycling episodes. For both cases, Tables \ref{tab:main-2015-processed} and \ref{tab:ep-2015-processed} contain labeled variables, facilitating the interpretation of the data.

```{r gss-processed-file-2015, echo=FALSE, cache=FALSE}
main_2015_processed <- gss_main_2015[1:10, 1:6]  %>%
  kable(
    format = "latex", 
    booktabs = TRUE, 
    longtable = TRUE, 
    align = c(rep("c", ncol(gss_m_2015[1:10, 1:6]))),
    caption = "\\label{tab:main-2015-processed}Visualization of the first ten lines and first six columns of the main file of the 2015 GSS.") %>%
  kable_styling(full_width = FALSE, 
                #latex_options=c("striped", "scale_down"), 
                latex_options=c("scale_down"),
                 font_size = 8) %>%
  footnote(general = c("Legend: `PUMFID`: record identification. `WGHT_PER`:  person weight. `SURVMNTH`: survey month of data collection. `AGEGR10`: age group of the respondent. `SEX`: sex of the respondent. `MARSTAT`: marital status of the respondent."), threeparttable = T)

@@

ep_2015_unprocessed <- gss_episodes_renamed |>
  filter(PUMFID == 10041 & YEAR == 2015) %>%
  dplyr::select(PUMFID, WGHT_EPI, act_label, DURATION, orig_label, dest_label, MODE) %>%
  kable(
    format = "latex", 
    booktabs = TRUE, 
    longtable = TRUE, 
    caption = "\\label{tab:ep-2015-processed}Visualization of active travel episodes of the record number `10041` for the 2015 GSS survey.", 
       col.names = c("PUMFID", "WGHT_EPI","Activity","Duration","Origin", "Destination", "Mode")) %>% 
  kable_styling(full_width = FALSE, 
                #latex_options=c("striped", "scale_down"), 
                latex_options=c("scale_down"),
                 font_size = 8) %>%
  footnote(general = c("Legend: `PUMFID`: record identification. `EPINO`: episode number. WGHT_EPI: episode's weight. TUI_01: activity code. DURATION: episode's duration. LOCATION: episode's location."), threeparttable = T)

@@

Considering GSS Cycles analyzed, we identified `r nrow(gss_episodes)` episodes that recorded active travel episodes, with trip duration ranging from 0 to 900 minutes, to twelve different destinations. `ActiveCA` includes all these episodes ready for analysis. Table \ref{tab:table-01} presents descriptive statistics on walking and cycling trips between 1986 and 2015, with measures of trip duration in minutes: maximum (max), mean, median, and minimum (min). The 1986 survey did not
include bicycle trips.

@@

`ActiveCA` also enables obtaining insights from the main processed files. Figure \ref{fig:figure-stress} present how the level of stress varied among respondents depending on their marital status in 2015. According to this plot, married respondents reported the highest level of stress, with 17% of possible cases.

@@

`ActiveCA` also provides a Jupyter Notebook containing a Python script that demonstrates how to read R data files (.rda) and convert them into Pandas DataFrames. This process allows users to work with and utilize the datasets available in `ActiveCA` within a Python project.

@@

This paper presents `ActiveCA`, an open data product that provides analysis-ready 
data from Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), and 29 (2015) 
of the GSS surveys on active travel in Canada. In the form of an `R` data package, 
{ActiveCA} was developed after collecting, cleaning, and processing the survey data, 
providing information on origins, destinations, and duration of active travel, as
well other information.

>>>>

---
title: "ActiveCA: Time Use Data from the General Social Survey of Canada to Study Active Travel"
author:
  - name: "Anon1"
    email: "Anon1@anon.org"
    affiliation: CODE
    footnote: 1
  - name: "Anon2"
    email: "Anon2@anon.org"
    affiliation: CODE
  - name: "Anon3"
    email: "Anon3@anon.org"
    affiliation: CODE
address:
  - code: CODE
    address: "Some school"
footnote:
  - code: 1
    text: "Corresponding Author"
abstract: 
  "This paper describes {ActiveCA}, an open data product with Canadian time use data. {ActiveCA} is an `R` data package that contains analysis-ready data related to active travel spanning almost 40 years, extracted from cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), and 29 (2015) of Canada's General Social Survey. Active travel is characterized by mode, with walking being part of every cycle and bicycling starting in 1992. The attributes of active trips are the types of locations of origins and destinations, the duration of trips, and episode weights for expanding the trips to population-wide estimates. Based on the year of the survey, a variety of locations are coded. In earlier cycles, these include home, work or school, and other's home, whereas in later cycles these are augmented with locations such as grocery stores, restaurants, outdoor destinations, and others. The geographical resolution includes the province and whether the episode was in an urban or rural setting."

@@

# Original main and Episode Files from 2015 to display table examples
gss_m_2015 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2015/gss-m.csv"))
gss_e_2015 <- read.csv(paste0(here(), "/data-raw/source-files/Time-use-2015/gss-e.csv"))

# Processed Main Files
data("gss_main_2015")
data("gss_main_2010")
data("gss_main_2005")
data("gss_main_1998")
data("gss_main_1992")
data("gss_main_1986")
```

@@

The Main File of the time use survey compiles a large array of aggregated data, 
summarizing the answers to the questionnaire that describe households and 
individuals, as well as derived variables that summarize the respondents' use of 
time use across different activities, locations, and social interactions. This 
file documents the time and duration that respondents allocate to each activity 
and location. The Main File provides a overview of daily routines and social 
dynamics, not focusing on individual activity episodes. Additionally, this file 
categorizes activities into bigger groups and subcategories, facilitating the 
data's analytical utility with additional metrics such as total transit time, time 
spent with household members, and counts of activities and episodes.

Table \ref{tab:main-2015-unprocessed} shows the first ten rows and first six variables of the GSS PUMF 2015 Main File (Cycle 29). Each row in the table correspond to a survey respondent, while the columns refer the following information: record identification (`PUMFID`), the person's weight (`WGHT_PER`), the month the survey data was collected (`SURVMNTH`), the respondent's age group (`AGEGR10`), the respondent's sex (`SEX`), and the respondent's marital status (`MARSTAT`).

The Main File of the 2015 GSS surveys includes a total of 17,390 respondents, representing `r sum(gss_m_2015$WGHT_PER)` individuals and 848 variables.  For discrete variables, Statistics Canada has assigned specific codes to the possible values, with each code accompanied by a label. For instance, in the case of the variable `SURVMNTH`, a value of `1` means `January 2016`, 2 means `February 2016`, 3 corresponds to `March 2016`, and so on.

@@

```{r gss-main-file-2015}
main_2015_unprocessed <- gss_m_2015[1:10, 1:6]  %>%
  kable(
    format = "latex", 
    booktabs = TRUE, 
    longtable = TRUE, 
    align = c(rep("c", ncol(gss_m_2015[1:10, 1:6]))),
    caption = "\\label{tab:main-2015-unprocessed}Visualization of the first ten lines and first six columns of the Main File of the 2015 GSS.") %>%
  kable_styling(full_width = FALSE, 
                #latex_options=c("striped", "scale_down"), 
                latex_options=c("scale_down"),
                 font_size = 8) %>%
  footnote(general = c("Legend: `PUMFID`: record identification. `WGHT_PER`:  person weight. `SURVMNTH`: survey month of data collection. `AGEGR10`: age group of the respondent. `SEX`: sex of the respondent. `MARSTAT`: marital status of the respondent."), threeparttable = T)

@@

The episode is a much bigger file that records detailed data for each activity 
episode reported by respondents. Each episode represents a single activity and its
duration, and the sum of all episodes throughout the day adds up to 24 hours. Each
entry in this file includes the start and end times of the activity, the duration,
location, and accompanying social context, informing when and where activities 
occurred and with whom. The focus of the Episode File is not on the characteristics
of the respondents, but on the characteristics of the activities, and the data 
are structured around the numerous activity instances that compose a day of the 
respondent. Although respondent-specific characteristics are not included within
the Episode File, it is possible to link the Main File and the Episode File by 
using a key present in both the Main and Episode Files.

Similar to Table \ref{tab:main-2015-unprocessed}, which displayed an example of the Main File structure, Table \ref{tab:ep-2015-unprocessed} presents the first seven episodes for the record identification number `10041` and some variables from the GSS PUMF 2015 Episode File (Cycle 29). Each row in the table corresponds to an episode associated with the specified record identification (`PUMFID = 10041`), including the episode's weight (`WGHT_EPI`), episode number (`EPINO`), activity code (`TUI_01`), episode duration (`DURATION`), and episode location (`LOCATION`).

In total, the Episode File of the 2015 GSS surveys contains 274,108 records, representing `r sum(gss_e_2015$WGHT_EPI)` episodes and 527 variables detailing the episodes. Similar to the Main File, Statistics Canada has created codes for the discrete variables, with each value corresponding to a label.

@@

For each selected cycle of the GSS surveys, we reviewed the Episode Files to identify episodes of movement that involved walking or cycling. This allowed us to select the activities immediately before and after the movement episode. After that, we labeled the code variables with their appropriate descriptions, identifying each origin and destination, mode of travel, time spent in the active trip, as well the province and urban classification of the episode.

@@

For each GSS Main File, we selected socioeconomic variables to help profile the individuals engaged in active trip episodes. We included key socioeconomic variables such as the respondent’s age group, sex, marital status, and number of children, among others. After this, as we did with the Episode Files, we labeled the code variables with their appropriate descriptions.

# How to use {ActiveCA}

This section presents some potential applications of the {ActiveCA} R package. In fact, we expect that the application of this package to extend beyond our pre-imagined range of uses.

To start the demonstration, it is first necessary to install the package. The code chunk below demonstrates how to install the {ActiveCA} package in R:

@@

After installing the {ActiveCA} R package, you can load it using the following code:

@@

For instance, if someone wants to calculate the percentage of respondents from the 2015 GSS survey with active travel episodes, it is necessary to account for the person's weights. In 2015, the weight variable is represented by `WGHT_PER.` The code chunk below demonstrates how to obtain the percentage of people with active travel episodes by age group. It uses the `dplyr`package [@dplyr2023] to manipulate the data.

The process begins by creating a dataset that sums the population by age group. Then, it joins the 2015 episodes with the 2015 Main File. Note that in both operations, the code sums the person weight variable (`WGHT_PER`) to obtain the correct population values. After this, a new dataset, `Active_percentage`, is created by merging both previous datasets. The percentage of active travel episodes by age group is calculated by dividing the total population by the population with active trip episodes, then multiplying by 100 and rounding with 2 decimal places.

@@

```{r percentages-act, echo=TRUE, out.width='80%'}
# Load the dplyr library for data manipulation
library('dplyr')

@@

# Combine total and active population 
#and calculate the percentage
Active_percentage <- Total_population %>% 
  left_join(Active_population, by = c("AGEGR10" = "AGEGR10")) %>%
  mutate(
    Percentage = round(Active_population/Total_population*100,2))

@@

Table \ref{tab:processed-obs} displays the total number of records processed for Main and Episode Files. For the Main Files, a total of `r sum(df_main_processed$Count)` records were processed, referring to all records from the Time Use Surveys from 1986 to 2015, that together represents more of `r sum(df_main_processed$Weighted)` respondents (`r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '2015',]$Weighted)` for 2015, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '2010',]$Weighted)` for 2010, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '2005',]$Weighted)` for 2005, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '1998',]$Weighted)` for 1998, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '1992',]$Weighted)` for 1992, and  for 2005, `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '1998',]$Weighted)` for 1998, and `r sum(df_main_processed[df_main_processed$Survey == 'Main' & df_main_processed$Year == '1992',]$Weighted)` for 1986). It also presents the total cases of active trips episodes identified. In total `r sum(df_ep_processed$Count)` records with register of active travel activity. Together, these records account for `r sum(df_ep_processed$Weighted)` episodes.

@@

table_df_processed <- df_processed %>%
  kable(format = "latex", booktabs = TRUE, longtable = TRUE, 
        caption = "\\label{tab:processed-obs}Total number and weighted sum of records processed.") %>%
  kable_styling(full_width = FALSE, font_size = 10) %>%
  add_header_above(c(" " = 2, "Observations" = ncol(df_processed) - 2)) %>%
  collapse_rows(columns = 1, valign = "top", latex_hline = "major")

@@

Table \ref{tab:main-2015-processed} shows the first ten rows and first six variables of the GSS PUMF 2015 Main File (Cycle 29), displayed in \ref{tab:main-2015-unprocessed} before our processing. Table \ref{tab:ep-2015-processed} presents the walking episodes for the record identification number `10041` from the GSS PUMF 2015 Episode File (Cycle 29), previously displayed in Table \ref{tab:ep-2015-unprocessed}. Only the unique active travel episode appear in Table \ref{tab:ep-2015-processed} since the records were filtered to select cases with walking or cycling episodes. For both cases, Tables \ref{tab:main-2015-processed} and \ref{tab:ep-2015-processed} contain labeled variables, facilitating the interpretation of the data.

```{r gss-processed-file-2015, echo=FALSE, cache=FALSE}
main_2015_processed <- gss_main_2015[1:10, 1:6]  %>%
  kable(
    format = "latex", 
    booktabs = TRUE, 
    longtable = TRUE, 
    align = c(rep("c", ncol(gss_m_2015[1:10, 1:6]))),
    caption = "\\label{tab:main-2015-processed}Visualization of the first ten lines and first six columns of the Main File of the 2015 GSS.") %>%
  kable_styling(full_width = FALSE, 
                #latex_options=c("striped", "scale_down"), 
                latex_options=c("scale_down"),
                 font_size = 8) %>%
  footnote(general = c("Legend: `PUMFID`: record identification. `WGHT_PER`:  person weight. `SURVMNTH`: survey month of data collection. `AGEGR10`: age group of the respondent. `SEX`: sex of the respondent. `MARSTAT`: marital status of the respondent."), threeparttable = T)

@@

ep_2015_unprocessed <- gss_episodes_renamed |>
  filter(PUMFID == 10041 & YEAR == 2015) %>%
  dplyr::select(PUMFID, WGHT_EPI, act_label, DURATION, orig_label, dest_label, MODE) %>%
  kable(
    format = "latex", 
    booktabs = TRUE, 
    longtable = TRUE, 
    caption = "\\label{tab:ep-2015-processed}Visualization of the active travel episode for the record number `10041` of the 2015 GSS survey.", 
       col.names = c("PUMFID", "WGHT_EPI","Activity","Duration","Origin", "Destination", "Mode")) %>% 
  kable_styling(full_width = FALSE, 
                #latex_options=c("striped", "scale_down"), 
                latex_options=c("scale_down"),
                 font_size = 8) %>%
  footnote(general = c("Legend: `PUMFID`: record identification. `EPINO`: episode number. WGHT_EPI: episode's weight. TUI_01: activity code. DURATION: episode's duration. LOCATION: episode's location."), threeparttable = T)

@@

Considering GSS Cycles analyzed, we identified `r nrow(gss_episodes)` episodes that recorded active travel episodes, with trip duration ranging from 0 to 900 minutes, to twelve different destinations. {ActiveCA} includes all these episodes ready for analysis. Table \ref{tab:table-01} presents descriptive statistics on walking and cycling trips between 1986 and 2015, with measures of trip duration in minutes. The 1986 survey did not include bicycle trips.

@@

{ActiveCA} also enables obtaining insights from the main processed files. Figure \ref{fig:figure-stress} present how the level of stress varied among respondents depending on their marital status in 2015. According to this plot, married respondents reported the highest level of stress, relating to feel stressed every day, with 17% of possible cases.

@@

{ActiveCA} also provides a Jupyter Notebook containing a Python script that demonstrates how to read R data files (.rda) and convert them into Pandas DataFrames. This process allows users to work with and utilize the datasets available in {ActiveCA} within a Python project.

@@

This paper presents {ActiveCA}, an open data product that provides analysis-ready 
data from Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24 (2010), and 29 (2015) 
of the GSS surveys on active travel in Canada. In the form of an `R` data package, 
{ActiveCA} was developed after collecting, cleaning, and processing the survey data, 
providing information on origins, destinations, and duration of active travel, as
well other information.
